---
title: "GGO parties codebook"
output: rmarkdown::html_vignette
author: James Hollway
date: "2025-09-28"
vignette: >
  %\VignetteIndexEntry{GGO parties codebook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(manytreaties)
```

## Release 1.0

This document provides a brief overview of the coding rationale for key variables 
in the dataset of states' relationships to international agreements 
provided in `manytreaties::parties$GGO`. 

Note that this dataset was constructed as a complement to datasets such as 
the International Environmental Agreements Database (`manytreaties::parties$IEADB`)
and the Design of Trade Agreements Database (`manytreaties::parties$DESTA`).
As such, it is neither complete in observations nor variables,
but offers some additional or alternative coding to other datasets.

Work on this dataset was supported by the Swiss National Science Foundation (SNSF)
[Grant Number 188976](https://data.snf.ch/grants/grant/188976): 
"Power and Networks and the Rate of Change in Institutional Complexes" (PANARCHIC).

Please direct all comments and suggestions to:

<center>
_James Hollway_

_International Relations/Political Science Department_

_Graduate Institute of International and Development Studies_

_Geneva, Switzerland_

_james.hollway@graduateinstitute.ch_
</center>

The variables are as follows:
`r paste(names(manytreaties::parties$GGO), collapse = ", ")`.

