---
title: "GGO agreements codebook"
output: rmarkdown::html_vignette
author: James Hollway
date: "2025-09-27"
vignette: >
  %\VignetteIndexEntry{GGO agreements codebook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(manytreaties)
```

## Release 1.0

This document provides a brief overview of the coding rationale for key variables in the list of episodes of independent states and state-like entities in the international system provided in `manytreaties::agreements$GGO`. 

Note that this dataset was constructed as a complement to datasets such as 
the International Environmental Agreements Database (`manytreaties::agreements$IEADB`)
and the Design of Trade Agreements Database (`manytreaties::agreements$DESTA`).
As such, it is incomplete in observations nor variables, 
yet offers some more specificity and some additional entries compared to such other datasets.

Work on this dataset was supported by the Swiss National Science Foundation (SNSF)
[Grant Number 188976](https://data.snf.ch/grants/grant/188976): 
"Power and Networks and the Rate of Change in Institutional Complexes" (PANARCHIC).

Please direct all comments and suggestions to:

<center>
_James Hollway_

_International Relations/Political Science Department_

_Graduate Institute of International and Development Studies_

_Geneva, Switzerland_

_james.hollway@graduateinstitute.ch_
</center>

The variables are as follows:
`r paste(names(manytreaties::agreements$GGO), collapse = ", ")`.

## Title, TitleAlt

This is the title of the treaty or agreement. 
We interpret an agreement broadly to include not only formal treaties,
such as treaties, conventions, or declarations,
but also other types of international agreements such as memoranda of understanding,
resolutions, protocols, and amendments.
We generally use the title of the agreement as listed on the treaty text itself,
or as listed in the source from which the treaty was identified.

In some situations an agreement might be known by several different titles.
These situations include:

- When treaties are recorded in different languages, or there are multiple valid versions of the treaty,
translation or transliterations can result in similar but different titles
- When treaties are recorded with different titles in different sources or databases (or versions of the database),
through translation or standardisation efforts
- When treaties are recorded with different titles by different parties to the treaty,
for example a bilateral agreement might be recorded under different names by each party
when they deposit it

In such situations, the shorter or more common title is generally preferred for the `Title` variable,
so long as it is unambiguous.
If only the longer version of the title mentions the parties, subject, or other distinguishing information,
then this title will be used as the `Title` variable.
Any alternative titles are included in the `TitleAlt` variable,
using commas to separate multiple alternative titles if necessary.

### Exclusion

While we strive to include a comprehensive list of international agreements,
including agrements that are not formal treaties, for instance,
but that can be filtered out by users if they wish (see `TypeAgree`),
there are some agreements that we do not include.
For the main part, it is because the information contained is not substantive,
but rather procedural or administrative.

- Exclude protocols of signature, accession, or confirmation (unless they come with substantive changes) -> these are dealt with in memberships data
- Exclude protocols of application or implementation, dealt with in entry into force date or (usually) as additional text
- Exclude amendments that are minor or technical, e.g. correcting typos, changing names, etc.

Note that exchanges of letters or similar documents that form part of a treaty are included.

### Differentiation

In some cases treaties with exactly the same title are in fact different treaties.
This might be because they were signed on different dates,
or because they were signed by different parties.
In other cases, treaties with different titles are actually the same treaty --
signed on the same date by the same parties.
We differentiated such entries as follows:

- confirmed duplicates were removed from the dataset
- confirmed non-duplicates were retained in the dataset but distinguished
from each other by adding distinguishing information, such as the year or subject,
to the treaty title(s).

### Deduplication

In other cases, treaties with very similar titles are actually the same treaties.
This is more challenging to determine.
We approach this deduplication task as follows:

- We first identify potential duplicates by searching for agreements that are concluded with the same date of signature
- From this list, we disregard those agreements that clearly concern different regions or subjects

With the list of remaining potential duplicates,
we then check the texts of the treaties themselves, if available,
or the sources from which they were identified.
We also use Google to search for the treaty title and date of signature,
to see if there are any sources that clarify whether these are the same or different treaties.
We use the following heuristics to assist with this process:

- Merge duplications representing multiple translations
- Similar treaty titles that refer to different articles, appendices, or annexes are retained separately
- Keep agreements referring to different parties
- Merge agreements that refer to the same parties but use different names for them, e.g. "the Government of X" and "the State of X"
- Merge agreements that are identifical except one mentions the parties and the other does not
- If we have other information that differs, such as different information for entry into force, this will cause a further check
- If Google returns (only) the same information, then this is a suggestion that these are duplicates
- We also note that some states or dyads have a history of concluding a set of agreements on related topics on the same day, e.g. China and Kazakhstan, and so this is also taken note of
- If there is a document combining multiple amendments on the same day to the same treaty, and also a listing separating out the amendments, then the combined amendment will be retained
- Merge duplications where substantive similarities and no treaty texts (or not for both) to indicate that they are different agreements

Note that agreements that are amendments or protocols to existing treaties are retained,
as these are often important in their own right.

## treatyID

A treatyID is a meaningful shorthand ID created from a combination of
elements extracted from the agreement title and date. 
The treatyID allows users to identify date, type and linkage.

### Bilateral treatyID

A bilateral treaty that is an agreement will have the following treatyID: "FRA-TON[DEZ]_1980A". 
This is a combination of the parties to the agreement (FRA-TON),
the activity described in the treaty,
which is here "Delimitation of Economic Zone (DEZ)",
with the year of signature (1980),
followed by the type (A = Agreement).

**Type** | **Pasting** | **Pasting**
--- | --- |--- 
Bilateral | FRA-TON[DEZ]_1980A | parties[activity]_year(type)
Bilateral + Protocol | 	RUS-USA[FKC]_1967E:RUS-USA[FKC]_1967A | parties[activity]_year(type):linkage

A bilateral treaty that is any other type than an agreement
(e.g. protocol, amendments) will have 
the treatyID under this format: "RUS-USA[FKC]_1967E:FKC_1965A". 
It is composed of the parties (RUS-USA),
the activity abbreviation from treaty title "Fishing for King Crab (FKC)", 
the year of signature of the amendment (1967), and
the type (E) which refers to Amendment. 
The linkage portion links the agreement to the treatyID its "mother" treaty 
(RUS-USA[FKC]_1967A).

### Multilateral treatyID

A multilateral treaty that is an agreement will have the following treatyID: "HSPDF_2005A".
The treatyID indicates the acronym (HSPDF),
the signature year of the agreement (2005),
and the type (A). 

**Type** | **Pasting** | **Pasting**
--- | --- |--- 
Multilateral | HSPDF_2005A | acronym_year(type)
Multilateral + Protocol | SFDP_2007E2:H08F_1992A | acronym_year(type and number):linkage

A multilateral treaty that is not an agreement will have this treatyID format: "SFDP_1992E2:H08F_2007". 
This represents the acronym (SFDP),
the signature year of the amendment (2007),
the type (E = Amendment),
the number of amendment (2),
and the linkage number (H08F_2007) of the "mother" treaty. 

**Type** | **Pasting** | **Pasting**
--- | --- |--- 
Known treaties | 	UNCLOS1982A | abbreviation, uID, type
Amendment of known treaties | J09H_1990E2:MARPOL1973A | acronym_uID, type(number):linkage

Famous multilateral treaties have a simplified treatyID with a known abbreviation. 
For example, the United Nations Convention On The Law Of The Sea
will have the following treatyID: "UNCLOC1982A". 
This is the known abbreviation (UNCLOS),
the signature year (1982) and the treaty type (A). 

Protocols or amendments of the known treaties will have this treatyID format: "J09H_1990E2:MARPOL1973A". 
It indicates the acronym (J09H),
the signature year of this specific amendment (1990),
the type of treaty (E),
its number (2), and the treatyID of the "mother" treaty (MARPOL197A).

