---
title: "GGO parties codebook"
output: rmarkdown::html_vignette
author: James Hollway
date: "2025-09-28"
vignette: >
  %\VignetteIndexEntry{GGO parties codebook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(manytreaties)
```

## Release 1.0

This document provides a brief overview of the coding rationale for key variables 
in the dataset of states' relationships to international agreements 
provided in `manytreaties::parties$GGO`. 

Note that this dataset was constructed as a complement to datasets such as 
the International Environmental Agreements Database (`manytreaties::parties$IEADB`)
and the Design of Trade Agreements Database (`manytreaties::parties$DESTA`).
As such, it is neither complete in observations nor variables,
but offers some additional or alternative coding to other datasets.

Work on this dataset was supported by the Swiss National Science Foundation (SNSF)
[Grant Number 188976](https://data.snf.ch/grants/grant/188976): 
"Power and Networks and the Rate of Change in Institutional Complexes" (PANARCHIC).

Please direct all comments and suggestions to:

<center>
_James Hollway_

_International Relations/Political Science Department_

_Graduate Institute of International and Development Studies_

_Geneva, Switzerland_

_james.hollway@graduateinstitute.ch_
</center>

The variables are as follows:
`r paste(names(manytreaties::parties$GGO), collapse = ", ")`.

## Identification: stateID, treatyID, Title

These are the identifying variables of this dataset.
The `stateID` is a standard three-letter country code as defined in `manystates::code_states()`.
The `Title` is the title of the agreement as listed on the treaty text itself,
or as listed in the source from which the treaty was identified.
The `treatyID` is a meaningful shorthand ID created from a combination of
elements extracted from the agreement title and date.
See `manytreaties::code_agreements()` for more details.

## Document dates: Begin, End, Signature, Force, Term

These are the dates when an agreement is deemed to have begun or ended.
Where more precise coding is available, we code also offer the date of signature,
the date when the agreement entered into force, and
the date when the agreement terminated.
These dates are drawn from `agreements$GGO`.

Dates are coded using the messydates system. 
This implements ISO’s extended date/time format. 
As such, some dates are only entered as a year or are annotated with a question mark if the source is uncertain. 
For more details see `{messydates}`.

Agreements that are known to still be in force at the time that the data was
collected or checked have an end date `9999-12-31`.
However, an `NA` in this field simply means that we could not verify the
status of the agreement at the time of coding.

