---
title: "GGO parties codebook"
output: rmarkdown::html_vignette
author: James Hollway
date: "2025-09-28"
vignette: >
  %\VignetteIndexEntry{GGO parties codebook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(manytreaties)
```

## Release 1.0

This document provides a brief overview of the coding rationale for key variables 
in the dataset of states' relationships to international agreements 
provided in `manytreaties::parties$GGO`. 

Note that this dataset was constructed as a complement to datasets such as 
the International Environmental Agreements Database (`manytreaties::parties$IEADB`)
and the Design of Trade Agreements Database (`manytreaties::parties$DESTA`).
As such, it is neither complete in observations nor variables,
but offers some additional or alternative coding to other datasets.

Work on this dataset was supported by the Swiss National Science Foundation (SNSF)
[Grant Number 188976](https://data.snf.ch/grants/grant/188976): 
"Power and Networks and the Rate of Change in Institutional Complexes" (PANARCHIC).

Please direct all comments and suggestions to:

<center>
_James Hollway_

_International Relations/Political Science Department_

_Graduate Institute of International and Development Studies_

_Geneva, Switzerland_

_james.hollway@graduateinstitute.ch_
</center>

The variables are as follows:
`r paste(names(manytreaties::parties$GGO), collapse = ", ")`.

## Identification: stateID, treatyID, Title

These are the identifying variables of this dataset.
The `stateID` is a standard three-letter country code as defined in `manystates::code_states()`.
The `Title` is the title of the agreement as listed on the treaty text itself,
or as listed in the source from which the treaty was identified.
The `treatyID` is a meaningful shorthand ID created from a combination of
elements extracted from the agreement title and date.
See `manytreaties::code_agreements()` for more details.

## Document dates: Begin, End, Signature, Force, Term

These are the dates when an agreement is deemed to have begun or ended.
Where more precise information is available, we code also the date of signature,
the date when the agreement entered into force, and
the date when the agreement terminated.
These dates are drawn from `agreements$GGO`.

Dates are coded using the messydates system. 
This implements ISO’s extended date/time format. 
As such, some dates are only entered as a year or are annotated with a question mark if the source is uncertain. 
For more details see `{messydates}`.

Agreements that are known to still be in force at the time that the data was
collected or checked have an end date `9999-12-31`.
However, an `NA` in this field simply means that we could not verify the
status of the agreement at the time of coding.

## Party dates: StateSignature, StateRatification, StateForce, StateEnd

These dates record dated actions that a state took with respect to an agreement,
to become a party to it, or to end its participation in it.
Its signature is the date that it first signed the agreement.
Its ratification is the date that it ratified, acceded to, or approved the 
agreement, depending on which the appropriate action is for the state,
agreement, and context.
Its force is the date that the agreement entered into force for that state,
typically being the date of ratification plus any period necessary
for the agreement to enter into force or other waiting period.
In some situations, the state terminates its relationship with the agreement
and this is recorded in the `StateEnd` variable.
Where a state terminates its relationship with an agreement, and it can rejoin,
this is recorded in a separate observation/row.
For the most part, this data is drawn from the same sources as we find
information about the agreements, their texts, and other metadata.
But in some cases we have drawn on other sources to verify or add this,
including through correspondence with responsible ministries of the parties.

## Further information: Coder, Comments

The `Coder` variable is a comma separated vector of the surnames of those who have added or verified data for each entry/observation. 
Where special conditions arise, the `Comments` variable offers a free text area for explanations or recording how the coding has changed from version to version. 
