---
title: "What Does Global Health Governance Look Like?"
author: "Jael Tan, Henrique Sposito, and Esther Peev"
date: "9 October 2024"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{What Does Global Health Governance Look Like?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(dplyr)
library(knitr)
library(kableExtra)
library(manydata)
library(manynet)
library(stringr)
```

## Health Agreements Data

`{manytreaties}` package contains data on health agreements and state memberships to these agreements.
  
  
### Agreements
  
  
The **agreements** datacube is an ensemble of data on international agreements. 
The datasets in this datacube provide an overview of international instruments 
that govern the global environment, trade, and health sphere. 

The HUGGO dataset in the `agreements` datacube
contains a handcoded data of the health-related agreements that builds upon data from the World Health Organization MiNDbank database and the Global Health and Human Rights database.
HUGGO improves the precision of dates, resolves any conflicts between these datasets,
corrects incorrect information,
and fills in any missing information that could be found.
  
  
When scraped automatically from the WHO MiNDBANK database online,
the latest date for agreements is 4313-12-31.
For this observation with the year '4313' in the WHO dataset,
"Human Rights Council Resolution A/HRC/RES/43/13 Mental Health and Human Rights",
the precise and accurate adoption date has been manually verified
and included in the HUGGO dataset as 2020-06-19.
With manual coding, the HUGGO dataset provides more precise and accurate data than what was obtained from webpage scraping.
  
  
Extending from data in the agreements datacube, the HUGGO dataset in the `parties` datacube contains hand-coded data of 39,000 observations on state memberships to international health instruments.
This is, to our knowledge, the first such dataset available.
  
  
```{r agreements, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
manytreaties::agreements$HUGGO[, 1:8] %>% 
  dplyr::filter(grepl("Health", Domain)) %>%
  dplyr::filter(messydates::year(Begin) == "2000") %>% 
  kable("html")

```


### Memberships

The **parties** datacube contains hand-coded data on states' memberships to instruments governing the global health sphere. 
The `HUGGO` dataset includes specific adoption/signature ('StateSignature'),
ratification ('StateRatification'), entry into force ('StateForce'),
and termination ('StateEnd') dates for each state during its membership to an agreement.
States that have predecessor or successor entities since 1945 are also identified in the Succession variable.

```{r mem, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
Domain <- manytreaties::agreements$HUGGO %>%
  dplyr::select(treatyID, Domain)
manytreaties::parties$HUGGO %>%
  dplyr::left_join(Domain, by = "treatyID") %>%
  dplyr::filter(grepl("Health", Domain)) %>%
  dplyr::filter(messydates::year(Begin) == "2005" & stateID == "CHE") %>% 
  kable("html")

```

Using the memberships data, we can explore the degree of overlap among states' membership to international health instruments.
For example, are members of the Pan-American Health Organization (PAHO) likely to be members of the same formal international health agreements?
The graph below shows that there is a high degree of overlap in the formal international health agreements joined by some of PAHO's largest state members 
(Argentina, Brazil, Canada, Mexico, Peru, Venezuela, and the United States) in the 2000s.

```{r memberships, warning=FALSE, message=FALSE, fig.align='center', fig.asp=1.2, fig.width=7}
pahoIDs <- c("ARG", "BRA", "CAN", "MEX", "PER", "VEN", "USA")
net <- manytreaties::parties$HUGGO %>%
  dplyr::left_join(Domain, by = "treatyID") %>%
  dplyr::filter(grepl("Health", Domain)) %>%
  dplyr::select(treatyID, stateID, Title, Begin) %>%
  dplyr::mutate(year = messydates::year(Begin)) %>%
  dplyr::filter(year > "1999" & year < "2010") %>%
  dplyr::filter(stateID %in% pahoIDs) %>%
  dplyr::distinct() %>%
  dplyr::select(treatyID, stateID) %>%
  as_tidygraph() %>%
  mutate(type = ifelse(stringr::str_detect(name, "[:digit:]{4}"), FALSE, TRUE))
max <- which(node_is_max(manynet::node_degree(net)))
net %>%
  mutate_ties(mem = ifelse(from %in% max, "all", "selective")) %>%
  graphr(layout = "hierarchy", edge_color = "mem")
```

For more information on how to interpret manyIDs, please read this [vignette](https://globalgov.github.io/manypkgs/articles/agreements.html) from `{manypkgs}` package.
For access to more data and information on our other "many" packages, please see [manydata](https://github.com/globalgov/manydata).
